Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2018-03-23 revised_prop_letter_ajhs.tex
revised_prop_letter_ajhs.tex \section{The Born Rule and Gleason's Theorem} Add new texts...

\NewDocumentCommand{\newInterval}{o o o o}{
 % \newcommand{#1}{\text{\wesa{{#2}}}}
 \newcommand{#1}{\ensuremath{\left[{#3}\right]}}
 % \newcommand{#1}{\ensuremath{I\left({#4}\right)}}
}
\newInterval[\imposs][impossible][0,0][0]
\newInterval[\likely][likely][\tfrac{1}{2},1][\tfrac{1}{2},1]
\newInterval[\unlikely][unlikely][0,\tfrac{1}{2}][0,\tfrac{1}{2}]
\newInterval[\necess][certain][1,1][1]
\newInterval[\unknown][unknown][0,1][0,1]
\newInterval[\midd][middle][\tfrac{1}{4},\tfrac{3}{4}][\tfrac{1}{4},\tfrac{3}{4}]

@@

\newcommand{\says}[3]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{#1}{#2 says: #3}\end{minipage}\end{framed}}
\newcommand{\amr}[1]{\says{green}{Amr}{#1}}
\newcommand{\yutsung}[1]{\says{purple}{Yu-Tsung}{#1}}
\newcommand{\gerardo}[1]{\says{OliveGreen}{Gerardo}{#1}}
\newcommand{\andy}[1]{\says{blue}{Andy}{#1}}
% \excludecomment{suggest}
\NewDocumentEnvironment{suggest}{m m O{}}{#1{TEXT \##2 would go here. #3}}{#1{TEXT \##2 end here.}}
\NewDocumentEnvironment{oldText}{m}{\clearpage{}#1{The following is old text.}}{#1{The old text ends here.}}
\NewDocumentEnvironment{newText}{m}{\columnbreak{}#1{The following is new text.}}{#1{The new text ends here.}}

@@

To that end, we construct a family of QIVPMs parameterized by
\emph{blurring} functions $f:\left[0,1\right]\rightarrow\mathscr{I}$ that
formalize how the real-valued probabilities are mapped to
intervals. We also introduce a notion of norm that quantifies the
distance between (pure or mixed) states.  The norm of a pure
state $\rho=\proj{\psi}$ is defined as usual by
$\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.  Similarly,
given any Hermitian operator~$A$, we consider the usual operator norm
$\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert
  =1}\left\Vert A\ket{\psi}\right\Vert $, which is also known as the
$2$-norm or the spectral
norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
Then, a finite-precision extension of Gleason's theorem can be stated as
follows.

@@

\begin{thm}[Finite-Precision Extension of the Gleason
  Theorem]\label{thm:Finite-precision-Gleason}Let $f$ be a blurring
  function and let $f \circ \muB_{\rho}$ be a QIVPM, where
  $\muB_{\rho}$ is the probability measure induced by the
  Born rule for a given state~$\rho$. Let $\alpha$ be the maximum
  length of intervals in its image. If a state $\rho'$ is consistent
  with $f\circ\muB_{\rho}$ on all events, i.e.,
  $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
  then the norm of their difference is bounded by $\alpha$, i.e.,
  $\left\Vert \rho-\rho'\right\Vert \le \alpha$.\end{thm}

The proof proceeds as follows. Given a state~$\rho'$ consistent with
$f\circ\muB_{\rho}$, we have
$\muB_{\rho'}\left(\proj{\psi}\right)\in
f\left(\muB_{\rho}\left(\proj{\psi}\right)\right)$ for any
one-dimensional projector $P=\proj{\psi}$. Since the maximum length of
the intervals in $\mathscr{I}$ is $\alpha$, it is also the upper bound
of the difference:
\[
\left|\muB_{\rho'}\left(\proj{\psi}\right)-\muB_{\rho}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|\le \alpha\,.
\]
Since $\rho-\rho'$ is Hermitian, $\max_{\left\Vert \psi\right\Vert =1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\cite{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\cite{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le \alpha$.

>>>>

\NewDocumentCommand{\newInterval}{o o o o}{
 \newcommand{#1}{\text{\wesa{{#2}}}}
 % \newcommand{#1}{\ensuremath{\left[{#3}\right]}}
 % \newcommand{#1}{\ensuremath{I\left({#4}\right)}}
}
\newInterval[\imposs][impossible][0,0][0]
\newInterval[\likely][likely][\tfrac{1}{2},1][\tfrac{1}{2},1]
\newInterval[\unlikely][unlikely][0,\tfrac{1}{2}][0,\tfrac{1}{2}]
\newInterval[\necess][certain][1,1][1]
\newInterval[\unknown][unknown][0,1][0,1]
\newInterval[\midd][middle][\tfrac{1}{4},\tfrac{3}{4}][\tfrac{1}{4},\tfrac{3}{4}]

@@

\newcommand{\says}[3]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{#1}{#2 says: #3}\end{minipage}\end{framed}}
\newcommand{\amr}[1]{\says{green}{Amr}{#1}}
\newcommand{\yutsung}[1]{\says{purple}{Yu-Tsung}{#1}}
\newcommand{\gerardo}[1]{\says{OliveGreen}{Gerardo}{#1}}
\newcommand{\andy}[1]{\says{blue}{Andy}{#1}}
% \excludecomment{suggest}
\NewDocumentEnvironment{suggest}{m m O{}}{#1{TEXT \##2 would go here. #3}}{#1{TEXT \##2 end here.}}
\NewDocumentEnvironment{compare}{}{\begin{widetext}\end{widetext}}{\begin{widetext}\end{widetext}}
\NewDocumentEnvironment{oldText}{m}{#1{The following is old text.}\noindent\begin{minipage}[t]{1\columnwidth}}{#1{The old text end here.}\end{minipage}\newpage}
\NewDocumentEnvironment{newText}{m}{\noindent\begin{minipage}[t]{1\columnwidth}#1{The following is new text.}}{\end{minipage}#1{The new text end here.}}

@@

\begin{compare}
\begin{oldText}{\yutsung}
To that end, we construct a family of QIVPMs parameterized by \emph{blurring}
functions $f:\left[0,1\right]\rightarrow\mathscr{I}$ that formalize
how the real-valued probabilities are mapped to intervals. We also
introduce a notion of norm that quantifies the distance between (pure
or mixed) states. The norm of a pure state $\rho=\proj{\psi}$ is
defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
Similarly, given any Hermitian operator~$A$, we consider the usual
operator norm $\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
Then, a finite-precision extension of Gleason's theorem can be stated
as follows.

@@

\begin{thm}[Finite-Precision Extension of the Gleason Theorem]\label{thm:Finite-precision-Gleason}Let
$f$ be a blurring function and let $f\circ\muB_{\rho}$ be a QIVPM,
where $\muB_{\rho}$ is the probability measure induced by the Born
rule for a given state~$\rho$. Let $\alpha$ be the maximum length
of intervals in its image. If a state $\rho'$ is consistent with
$f\circ\muB_{\rho}$ on all events, i.e., $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
then the norm of their difference is bounded by $\alpha$, i.e., $\left\Vert \rho-\rho'\right\Vert \le\alpha$.\end{thm}
\end{oldText}

\begin{newText}{\yutsung}
To that end, we construct a family of QIVPMs parameterized by \emph{blurring}
functions. A blurring function $f:\left[0,1\right]\rightarrow\mathscr{I}$
maps every real-valued probability $x\in\left[0,1\right]$ to a set
of intervals $f\left(x\right)=\left[\ell,r\right]$ containing $x$,
where $\left[0,1\right]$ denotes the set of real-valued probabilities
which should not be considered as the interval-valued probability
$\unknown$. We also introduce a notion of norm that quantifies the
distance between (pure or mixed) states. The norm of a pure state
$\rho=\proj{\psi}$ is defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
Similarly, given any Hermitian operator~$A$, we consider the usual
operator norm $\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
the norm is the maximum absolute value of its eigenvalue. Then, a
finite-precision extension of Gleason's theorem can be stated as follows.

\begin{thm}[Finite-Precision Extension of the Gleason Theorem]\label{thm:Finite-precision-Gleason-1}Let
$f:\left[0,1\right]\rightarrow\mathscr{I}$ be a blurring function
and let $f\circ\muB_{\rho}$ be a QIVPM, where $\muB_{\rho}$ is the
probability measure induced by the Born rule for a given state~$\rho$.
Let $\alpha$ be the maximum length of intervals in $\mathscr{I}$.
If a state $\rho'$ is consistent with $f\circ\muB_{\rho}$ on all
events, i.e., $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
then the norm of their difference is bounded by $\alpha$, i.e., $\left\Vert \rho-\rho'\right\Vert \le\alpha$.\end{thm}
\end{newText}
\end{compare}

The proof proceeds as follows. Given a state~$\rho'$ consistent
with $f\circ\muB_{\rho}$, we have $\muB_{\rho'}\left(\proj{\psi}\right)\in f\left(\muB_{\rho}\left(\proj{\psi}\right)\right)$
for any one-dimensional projector $P=\proj{\psi}$. Since the maximum
length of the intervals in $\mathscr{I}$ is $\alpha$, it is also
the upper bound of the difference: 
\[
\left|\muB_{\rho'}\left(\proj{\psi}\right)-\muB_{\rho}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|\le\alpha\,.
\]
Since $\rho-\rho'$ is Hermitian, $\max_{\left\Vert \psi\right\Vert =1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\cite{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\cite{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le\alpha$.
